---
title: "Activity Monitoring Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Introduction
It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This report makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

Code for reading in the dataset and/or processing the data

```{r readdata, echo=TRUE}
activity <- read.csv("data/activity.csv", stringsAsFactors = FALSE)
```

Histogram of the total number of steps taken each day

```{r histogram, echo=TRUE}
activity <- activity[!is.na(activity$steps),]
activity$steps <- as.numeric(activity$steps)
activity$date <- as.Date(activity$date, "%Y-%m-%d")
activity_sums <- aggregate(steps ~ date, activity, sum)
hist(activity_sums$steps, breaks = 5, main = "Frequency of Steps Per Day", xlab = "Steps Per Day", ylab = "Frequency", col = "red")
```

ggplot(activity_sums, aes(x = activity_sums$date, y = activity_sums$steps)) + geom_bar(stat="identity", fill="white", colour="darkgreen") + xlab("Date") + ylab("Steps") + ggtitle("Steps by Date")


Mean and median number of steps taken each day

Time series plot of the average number of steps taken

The 5-minute interval that, on average, contains the maximum number of steps

Code to describe and show a strategy for imputing missing data

Histogram of the total number of steps taken each day after missing values are imputed

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

All of the R code needed to reproduce the results (numbers, plots, etc.) in the report